\documentclass{vgtc}                % final (conference style)

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amsmath,amssymb}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}

\vgtcinsertpkg
\nocopyrightspace

%% Title
\title{AgentFlow: An Investigative Visual Analytics Platform for
Reconstructing\\ the TenantThread Embargo Breach (VAST Challenge 2026, MC-1)}

%% Authors: three blocks (left / center / right), e-mails at the top
\author{Jyns Arturo Ord\'o\~nez Del Carpio\\%
        \scriptsize Universidad de Ingenier\'ia y Tecnolog\'ia (UTEC), Lima, Peru\\%
        \scriptsize jyns.ordonez@utec.edu.pe %
\and Bihonda Epiro\\%
        \scriptsize Universidad de Ingenier\'ia y Tecnolog\'ia (UTEC), Lima, Peru\\%
        \scriptsize bihonda.epiro@utec.edu.pe %
\and Cesar Pajuelo\\%
        \scriptsize Universidad de Ingenier\'ia y Tecnolog\'ia (UTEC), Lima, Peru\\%
        \scriptsize cesar.pajuelo@utec.edu.pe}

%% Abstract
\abstract{
Mini-Challenge~1 of the VAST Challenge 2026 asks how embargoed
information about the \textit{Project HarborCrest} merger
(TenantThread--CivicLoom) reached the public FleX platform
$\sim$35 minutes before the 18:00, June~5, 2046 embargo lift,
given the communication log of seven autonomous agents. We
present \textbf{AgentFlow}, a forensic investigation console ---
deliberately not a dashboard --- built with D3.js~v7 on a
reproducible Python feature-engineering pipeline. The
investigation is condensed into two linked views: the
\emph{Behavioral Drift Braid}, which plots each agent as a thread
on an ordinal covert-to-public posture ladder composited with
expected baselines, covert-coordination arcs, compliance gates,
the causal chain E01--E12, and a leading-indicator seismograph
with pre-crisis $\mu{+}1\sigma$/$\mu{+}2\sigma$ thresholds; and
the \emph{Hypothesis Balance}, a transparent evidence model whose
twelve categories, weighted with diminishing returns
$w_c=b_c(1-e^{-n_c/s_c})$, physically tilt a scale between
\textsc{intentional} and \textsc{systemic failure}. Every signal
is a documented arithmetic formula, auditable down to the
original message from a persistent Evidence Inspector. Over the
912 messages the balance tilts 25.3 vs.\ 14.2 toward a breach
with a deliberate core, executed by the two agents who dominated
the covert channel and preceded by early indicators detectable
from May~22.%
}

\keywords{Visual analytics, behavioral anomaly detection,
communication forensics, evidence weighting, VAST Challenge.}

\begin{document}

\firstsection{Introduction}

\maketitle

TenantThread secretly negotiates its acquisition by CivicLoom
under an embargo until 18:00, June~5, 2046. Seven autonomous
agents (legal, social media, quality, PR, two interns, and a
compliance agent, \textit{The Judge}) exchange 912 messages over
six channels of decreasing traceability; the embargoed fact
surfaces publicly 35 minutes early. The analytic difficulty is
threefold: evidence is scattered across messages, private
reasoning (\texttt{internal\_state}) and 77 public posts, sampled
irregularly (daily, then hourly on the crisis day); no single
message constitutes the breach --- the signal is a distributed
gradient of covert coordination, boundary testing and incremental
disclosure via ``defensible'' governance language; and the
central question (deliberate leak vs.\ systemic failure) is a
judgment call. Our contributions: (1)~a fully documented pipeline
in which every score is an auditable formula; (2)~the
\emph{Behavioral Drift Braid}, turning behavioral drift into
literal geometry; (3)~a composite \emph{seismograph} with
statistically grounded early-warning thresholds; (4)~the
\emph{Hypothesis Balance}, an explicit evidence-weighting model
that supports --- not replaces --- the analyst's judgment.

\section{Data and Feature Engineering}

One script derives seven JSON datasets (messages, rounds,
network, actors, posts, events, meta). \emph{No score is a black
box.} Five transparent lexicon families flag both public content
(\texttt{nlp}: what was \emph{said}) and private reasoning
(\texttt{nlp\_inner}: what was \emph{known}): \textbf{merger}
(the embargoed subject: \texttt{civicloom}, \texttt{harborcrest},
\texttt{definitive agreement}), \textbf{embargo} (awareness:
\texttt{6 pm}, \texttt{confidential}, \texttt{material
nonpublic}), \textbf{execution} (the act: \texttt{executing},
\texttt{going live}), \textbf{compliance} (rationalization:
\texttt{10b-5}, \texttt{defensible}), and \textbf{governance} ---
the case-specific insight: the breach leaked through individually
defensible ``governance clarifications'' (\texttt{audit},
\texttt{consent management}, \texttt{access control}), not blunt
merger statements. Message sensitivity is
\begin{equation}
\label{eq:sens}
S(m) = 3.0\,|M_m| + 2.0\,|E_m| + 1.5\,|G_m| + 1.0\,|X_m|
\end{equation}
over the merger/embargo/governance/execution term counts, weights
decreasing with legal gravity of each family. An \emph{embargo
violation} is the strict conjunction $V(m)=$ (public $\wedge$
merger language $\wedge$ $t_m<t_{\text{embargo}}$); 16 messages
qualify. Each public post $p$ receives a leak likelihood
\begin{align}
\label{eq:leak}
LL(p) ={}& 0.45\,[|M_p|{>}0] + 0.20\,[t_p{<}t_{\text{emb}}]
        + 0.15\,[|X_p|{>}0] \nonumber\\
       &+ 0.10\,[\text{anonymous}] + 0.10\,\min(C(p),10)/10,
\end{align}
where the coordination score $C(p)$ counts covert
\texttt{side\_huddle} messages in the prior 90 minutes --- the
signature of a \emph{primed} act. Betweenness (Brandes) is
computed \textbf{only over private/covert edges}: the team is a
broadcast clique, so the global metric is trivially zero;
restricted, it reveals who bridges reserved conversations (legal:
0.36, maximum). \emph{Brokerage} $B(a)=$ (covert+private messages
sent) $\times$ (public posts authored), normalized, finds the
internal$\rightarrow$public conduit (social media 1.00, legal
0.78). Behavioral deviation is the normalized $L_1$ distance
$D(a)=\frac{1}{2}\sum_c|\pi^{\text{crisis}}_a(c)-\pi^{\text{base}}_a(c)|$
between crisis-day and pre-crisis channel mixes, plus
\texttt{new\_channels} never used in baseline.

\section{Visual Design}

\textbf{Rationale.} The layout is a forensic case file: filter
rail, central canvas, persistent \emph{Evidence Inspector}. Any
element loads its messages with matched terms highlighted and
internal reasoning visible --- visual abstraction to textual
proof in one click. A single \texttt{filtered()} state (actors,
channels, evidence flags, time brush, search) drives every view;
filters have all/none toggles and one-click Reset; searching
draws a per-round sparkline of term mentions. Time is ordinal by
round so the nine decisive hours are not crushed by two calm
weeks. Bilingual UI (EN/ES), light/dark themes.

\textbf{Behavioral Drift Braid.} Each agent is a thread on an
ordinal posture ladder $\Pi=\{$monitored huddle$\,{=}\,0$, 1:1$\,{=}\,1$,
side-huddle$\,{=}\,2$, official$\,{=}\,3$, personal$\,{=}\,4$,
anonymous$\,{=}\,5\}$; its per-round position is the weighted mean
\begin{equation}
\label{eq:posture}
y_a(r) = \frac{\sum_{m} w_m\,\Pi(c_m)}{\sum_{m} w_m},
\qquad w_m=3 \text{ if public, else } 1,
\end{equation}
so exposure visibly pulls the thread; crossing the red
\emph{enforcement perimeter} means content went public.
Composited layers: dotted pre-crisis baselines ($\mu\pm\sigma$ of
$y_a$: expected behavior), purple covert-coordination arcs, Judge
gates (green~=~held; red~=~a violation followed within
$[r,r{+}2]$) with the auto-detected \textsc{judge silent} band,
the causal chain E01--E12 with decision arrows (including the red
``ceiling warning ignored'' arrow E09$\rightarrow$E10), and one
triangle per public post, red-rimmed if violating. A \emph{Key
Questions} panel reconfigures the braid per challenge task with
linked evidence.

\textbf{Seismograph.} Each round is scored
\begin{equation}
\label{eq:anomaly}
A(r) = 0.4\,\Sigma S_r + 1.0\,n^{\text{covert}}_r
     + 3.0\,n^{\text{boundary}}_r + 6.0\,n^{\text{viol}}_r
\end{equation}
against dashed pre-crisis $\mu{+}1\sigma$ (early warning) and
$\mu{+}2\sigma$ (alert) thresholds, next to the \$TTHR share
price and per-round covert/boundary/violation tick rows.

\section{The Hypothesis Balance}

Each evidence category is a \emph{physical weight} dropping into
the red (\textsc{intentional}) or blue (\textsc{systemic
failure}) pan of an SVG scale animated by a force simulation. The
effective weight of category $c$ with $n_c$ matching messages
saturates:
\begin{equation}
\label{eq:sat}
w_c = b_c\big(1 - e^{-n_c/s_c}\big),
\end{equation}
so the first hit counts most and \emph{volume cannot drown
signal}: one attributable official-channel violation ($b{=}8$)
outweighs dozens of routine governance mentions ($b{=}2$).
Table~\ref{tab:categories} lists all twelve categories with their
exact assignment rules --- also shown in-UI as transparency
tooltips (why it matters, routing rule, $b_c$, $n_c$, pole);
clicking a row loads its messages. The structural concentration
term encodes that concentrated breaches indicate design:
$w_{\text{conc}} = 6.5\cdot\sum_{a\in\text{top-2}}v_a/\sum_a v_a$
over per-agent violation counts $v_a$. Categories map with signed
multipliers to the three hypotheses (A~deliberate,
B~coordination failure, C~emergent with deliberate core);
$\text{net}_h=\max(0,\Sigma\text{FOR}-\Sigma\text{AGAINST})$,
$\text{conf}_h=\text{net}_h/\Sigma\text{net}$, closing as
\textsc{confirmed} at $\geq$45\% with a $\geq$12-point lead or
\textsc{rejected} at $\leq$15\%. The beam tilt
$\theta=0.26\tanh((W_{\text{sys}}-W_{\text{int}})/7)$ keeps
extreme imbalances readable; a verdict banner restates the tilt
in plain language under every filter. The balance shows the
\emph{balance of evidence}, never a machine conclusion.

\begin{table}[htb]
\caption{Evidence categories: base weight, saturation, rule.}
\label{tab:categories}
\centering
\scriptsize
\begin{tabular}{p{1.9cm}ccp{3.2cm}}
\toprule
\textbf{Category} & $b_c$ & $s_c$ & \textbf{Assignment rule} \\
\midrule
\multicolumn{4}{l}{\emph{INTENTIONAL pole}} \\
Attributable leak & 8 & 2 & public $\wedge$ violation $\wedge$ official channel \\
Anonymous pre-seeding & 5 & 2 & anonymous $\wedge$ merger $\wedge$ before embargo \\
Covert coordination & 4 & 20 & merger/embargo in \texttt{side\_huddle} \\
Breach concentration & $6.5\times$ & --- & top-2 agents' share of violations \\
Breach timing cluster & 6 & 1 & violation $\wedge$ crisis-day round \\
Channel escalation & 5 & 1 & same agent breaches via covert \emph{and} personal/anonymous \\
\midrule
\multicolumn{4}{l}{\emph{SYSTEMIC FAILURE pole}} \\
Boundary testing & 3 & 8 & public $\wedge$ merger/embargo $\wedge$ $\neg$violation \\
Uncontrolled leak & 4 & 6 & violation $\wedge$ non-official channel \\
Defensive clarification & 2 & 3 & public $\wedge$ compliance $\wedge$ $\neg$violation \\
Late-installed control & 3 & fixed & Judge deployed $\sim$1 week before breach (E04) \\
Unenforced warning & 1.5 & fixed & E09 issued without coercive power \\
Channel reinforcement & 3 & --- & merger live on all 3 public channels at once \\
\bottomrule
\end{tabular}
\end{table}

\section{Answers to the MC-1 Tasks}

\textbf{Q1 --- Events, relationships, decision points.} The chain
reads on the braid: CEO seeds ``strategic developments''
privately (May~17, E01) $\to$ covert \texttt{side\_huddle}
activated, merger briefing held there (May~22, E02--E03) $\to$
Judge assigned only after a public faux pas (May~30, E04) $\to$
SaltWind expos\'e pressure (May~31/Jun~4, E05--E06) $\to$ June~5:
anonymous post references the deal (09:49, E07), boundary tests
escalate (E08), the 15:08 ceiling warning (E09) is ignored $\to$
Legal posts on its personal account at 17:25 (E10), Social-Media
amplifies (E11), breach confirmed before 18:00 (E12). Key
decisions: activating the covert channel, briefing there,
ignoring E09, posting.

\textbf{Q2 --- What let the post past enforcement.} (1)~Late
control: gates appear only from May~30, after coordination had
gone covert; (2)~the silent gap: zero compliance output during
the crisis morning when the first violations occurred;
(3)~advisory-only power: E09 is a ruling, not a block;
(4)~scope: enforcement watched the official account while the
breach travelled personal/anonymous channels --- exactly where
the Legal and Social-Media threads end the day.

\textbf{Q3 --- Typical vs.\ breach behavior.} Two pre-crisis
weeks of flat threads in monitored zones define normality; on
June~5 the Legal thread dives across the perimeter: $D=0.31$, 176
crisis-day messages vs.\ a 90-message baseline, 12 anonymous
posts, and \emph{two never-before-used public channels}.
Social-Media mirrors the dive; all other agents stay near
baseline, isolating the anomaly to the same pair that dominates
brokerage.

\textbf{Q4 --- Leading indicators.} Yes: the seismograph exceeds
$\mu{+}1\sigma$ on May~22 (covert activation) and May~29 (public
near-incident) and crosses $\mu{+}2\sigma$ hours before the
breach; anonymous pre-seeding starts at 09:49, eight hours before
confirmation.

\textbf{Q5 --- Prior deviations.} May~22: senior agents move
merger talk off the monitored channel (first posture drop, first
purple arcs). May~29: @Elena's public post skirts merger language
--- the faux pas that triggered the Judge's assignment.
May~31--June~4: governance-framed clarifications rise --- each
locally defensible, jointly disclosing.

\textbf{Q6 --- Similar episodes.} The pattern (covert alignment
$\to$ public boundary test) rehearses twice: May~22--23 (covert
cluster + merger-adjacent post) and May~29--30 (near-incident +
defensive clarifications); brushing to May makes both visually
isomorphic to June~5.

\textbf{Q7 --- Why no earlier action.} Green gates were read as
success (false confidence); each boundary test was individually
defensible, so rule-based alerts never fired; the only escalation
was advisory. The system punished blunt violations but was blind
to the \emph{gradient} --- what the composite score makes
visible.

\textbf{Verdict.} The balance tilts \textsc{intentional} 25.3
vs.\ 14.2: \emph{emergent behavior with a deliberate core} ---
unmonitored channels and late controls enabled the leak; a
deliberate, attributable act completed it.

\section{Conclusion}

Three design decisions proved decisive: the ordinal posture
ladder with $\times 3$ public weighting, betweenness restricted
to private edges, and exponential evidence saturation --- rare
conclusive evidence beats routine volume. Base weights remain
editorial (visible and defended, not learned) and the verdict is
filter-sensitive by design. AgentFlow does not decide for the
analyst; it makes the analyst's judgment \emph{defensible}.

\end{document}
